---
output: html_document
---
## Part 2. Basic inferential data analysis

Load data.
```{r}
library(datasets)
data(ToothGrowth)
```

Let's explore the data.

```{r}
head(ToothGrowth)
str(ToothGrowth)
summary(ToothGrowth)
table(ToothGrowth$supp, ToothGrowth$dose)
```

We describe that ToothGrowth is a data frame with three variables - length in millimeter, OJ and VC supplements, and doses (0.5, 1,0 and 2.0). Two types of supplements are administered to two groups of 30 subjects each to promote tooth growth. Each supplement is given in 0.5, 1.0 and 2.0 doses to ten subjects each.

```{r}
library(ggplot2)
ggplot(ToothGrowth, aes(len, dose)) + geom_point(aes(color = supp), size = 4, alpha = 1/2) + labs(title = "Tooth Growth") + labs(x = "Length", y = "Dose")
```

* It appears that increasing dose (0.5 -> 1.0 -> 2.0) results in longer growth.
* Which supplement is better?
* We will not assumt that higher than 2.0 doses results in even longer growth.

### Relationship between dose and length

H0: The mean lengths corresponding to different doses are the same or in other words there is no relationship between the dose and length.

HA: The mean lengths corresponding to different doses are different and there is a relationship between the dose and length.

The groups have the same variance.

Compare 0.5 and 1.0 dosages

```{r}
TG0.5and1.0 = subset(ToothGrowth, dose %in% c(0.5, 1.0))
t.test(len ~ dose, paired = FALSE, var.equal = TRUE, data = TG0.5and1.0)
```

The test suggest that we reject the null hypothesis in favor of the alternative based on p-value = 1.266e-07. The difference in means (-9.13) would only happen by chance 1 time in 1.266 * 10^7 experiments. Thus, we conclude that the difference in means is not a chance, but it is due to a real effect of higher dosage.

Comparing  1.0 and 2.0 dosages yields a similar result.

```{r}
TG1.0and2.0 = subset(ToothGrowth, dose %in% c(1.0, 2.0))
t.test(len ~ dose, paired = FALSE, var.equal = TRUE, data = TG1.0and2.0)
```

### Which supplement is better?

The two supplements will have differenct variances.

Compare OJ and VC

```{r}
t.test(len ~ supp, paired = FALSE, var.equal = FALSE, data = ToothGrowth)
```

OJ is better and the difference ~4 mm lies within the 95% confidence interval.