##Q1
Suppose I define the following function in R
```{r}
cube <- function(x, n) {
        x^3
}
```
What is the result of running
```{r eval=F}
cube(3)
```
in R after defining this function?

##A1
```{r}
cube(3)
```

##Q2
The following code will produce a warning in R.
```{r eval=F}
x <- 1:10
if(x > 5) {
        x <- 0
}
```
Why?

##A2
'x' is a vector of length 10 and 'if' can only test a single logical statement.

##Q3
Consider the following function
```{r}
f <- function(x) {
        g <- function(y) {
                y + z
        }
        z <- 4
        x + g(x)
}
```
If I then run in R
```{r eval=F}
z <- 10
f(3)
```
What value is returned?

##A3
```{r}
z <- 10
f(3)
```

##Q4
Consider the following expression:
```{r}
x <- 5
y <- if(x < 3) {
        NA
} else {
        10
}
```
What is the value of 'y' after evaluating this expression?

##A4
```{r}
y
```

##Q5
Consider the following R function
```{r}
h <- function(x, y = NULL, d = 3L) {
        z <- cbind(x, d)
        if(!is.null(y))
                z <- z + y
        else
                z <- z + f
        g <- x + y / z
        if(d == 3L)
                return(g)
        g <- g + 10
        g
}
```
Which symbol in the above function is a free variable?

##A5
f

##Q6
What is an environment in R?

##A6
a collection of symbol/value pairs 

##Q7
The R language uses what type of scoping rule for resolving free variables?

##A7
lexical scoping 

##Q8
How are free variables in R functions resolved?

##A8
The values of free variables are searched for in the environment in which the function was defined

##Q9
What is one of the consequences of the scoping rules used in R?

##A9
All objects must be stored in memory 

##Q10
In R, what is the parent frame?

##A10
It is the environment in which a function was called

