##Question 1
R was developed by statisticians working at

##A1
The University of Auckland 

##Question 2
The definition of free software consists of four freedoms (freedoms 0 through 3). Which of the following is NOT one of the freedoms that are part of the definition?

##A2
The freedom to sell the software for any price. 

##Question 3
In R the following are all atomic data types EXCEPT 

##A3
array 

##Question 4
If I execute the expression x <- 4L in R, what is the class of the object 'x' as determined by the 'class()' function?

##A4
```{r}
x <- 4L
class(x)
```

##Question 5
What is the class of the object defined by the expression x <- c(4, "a", TRUE)?

##A5
```{r}
x <- c(4, "a", TRUE)
class(x)
```

##Question 6
If I have two vectors x <- c(1,3, 5) and y <- c(3, 2, 10), what is produced by the expression cbind(x, y)?

##A6
```{r}
x <- c(1, 3, 5)
y <- c(3, 2, 10)
cbind(x, y)
```

##Question 7
A key property of vectors in R is that

##A7
elements of a vector all must be of the same class 

##Question 8
Suppose I have a list defined as x <- list(2, "a", "b", TRUE). What does x[[1]] give me?

##A8
```{r}
x <- list(2, "a", "b", TRUE)
x[[1]]
class(x[[1]])
```

##Question 9
Suppose I have a vector x <- 1:4 and a vector y <- 2. What is produced by the expression x + y?

##A9
```{r}
x <- 1:4
y <- 2
x + y
class(x + y)
```

##Question 10
Suppose I have a vector x <- c(3, 5, 1, 10, 12, 6) and I want to set all elements of this vector that are less than 6 to be equal to zero. What R code achieves this?

##A10
```{r}
x <- c(3, 5, 1, 10, 12, 6)
x[x < 6] <- 0 
x
```

##Question 11
In the dataset provided for this Quiz, what are the column names of the dataset?

##A11
```{r}
d <- read.csv("hw1_data.csv")
names(d)
```

##Question 12
Extract the first 2 rows of the data frame and print them to the console. What does the output look like?

##A12
```{r}
d[1:2,]
```

##Question 13
How many observations (i.e. rows) are in this data frame?

##A13
```{r}
nrow(d)
```

##Question 14
Extract the last 2 rows of the data frame and print them to the console. What does the output look like?

##A14
```{r}
tail(d, n = 2)
```

##Question 15
What is the value of Ozone in the 47th row?

##A15
```{r}
d[47, "Ozone"]
```

##Question 16
How many missing values are in the Ozone column of this data frame?

##A16
```{r}
sum(is.na(d$Ozone))
```

##Question 17
What is the mean of the Ozone column in this dataset? Exclude missing values (coded as NA) from this calculation.

##A17
```{r}
mean(d[!is.na(d$Ozone), "Ozone"])
```

##Question 18
Extract the subset of rows of the data frame where Ozone values are above 31 and Temp values are above 90. What is the mean of Solar.R in this subset?

##A18
```{r}
i <- d$Ozone > 31 & d$Temp > 90
mean(d[i, "Solar.R"], na.rm = T)
```

##Question 19
What is the mean of "Temp" when "Month" is equal to 6? 

##A19
```{r}
mean(d[d$Month == 6, "Temp"])
```

##Question 20
What was the maximum ozone value in the month of May (i.e. Month = 5)?

##A20
```{r}
max(d[d$Month == 5, "Ozone"], na.rm = T)
```